# Regression and Model validation

*Describe the work you have done this week and summarize your learning.*

- Describe your work and results clearly. 
- Assume the reader has an introductory course level understanding of writing and reading R code as well as statistical methods
- Assume the reader has no previous knowledge of your data or the more advanced methods you are using  

During the second week of this course, we have been delving deeper into R and statistics. We learned about regression models and the application of R in statistical modeling. The datacamp exercises along with the two embedded videos provided good background on the topics. Chapter three from ["An Introduction to Statistical Learning with Applications in R"](http://www-bcf.usc.edu/~gareth/ISL/) covered in-depth topics in linear regression. 

## RStudio Exercise 2
### Data Wrangling
After going through the study materials, I attempted the RStudio exercise. The first part of the exercise was related to data wrangling where a subset of table was generated from a table with raw data (observations). The R script used to create the table can be found [here](https://github.com/kisun/IODS-project/tree/master/data/create_learning2014.R).

### Analysis
The R script for this part is available [here](https://github.com/kisun/IODS-project/tree/master/data/analysis.R).
The data used in this exercise comes from an international survey of approaches to learning conducted by Kimmo Vehkalahti. The survey was funded by Teachers' Academy funding (2013-2015) and the data was collected during December 2014 and January 2015. The survey was conducted in Finland with an aim to understand the relationship between learning approaches and students' achievements in an introductory statistics course. A total of 183 individuals were included in the survey where the students were assessed for three different studying approaches - surface approach, deep approach and strategic approach. Additional details about the survey can be found [here](http://www.slideshare.net/kimmovehkalahti/the-relationship-between-learning-approaches-and-students-achievements-in-an-introductory-statistics-course-in-finland
). After preprocessing in data wrangling step, we read the data into R and applied regression models.


```{r}
lrn2014<-read.table("data/learning2014.csv")
str(lrn2014)
dim(lrn2014)
```
The final table used for analysis consist data for seven different variables and 166 individuals (see above). Among the variables, gender is a factor variable, age and points are integers whereas attitude, deep, stra and surf variables include float values.

```{r, fig.width=11, fig.height=9, echo=FALSE}
library(GGally)
library(ggplot2)
```
```{r, fig.width=11, fig.height=9}
plot_lrn2014<-ggpairs(lrn2014, mapping = aes(col=gender, alpha = 0.3), lower=list(combo = wrap ("facethist", bins = 20)))
plot_lrn2014
```
The graphical overview of the data is shown above. The number of females (n=110) in this survey is almost two times the number of males (n=56).It seems that the females have comparatively more diverese age groups than male counterparts. Here, the overall goal of the survey is to identify how age of the students, attitude towards learning and three different learning methods are contributing towards the final points. In general, attitude towards learning has the highest impact for overall outcome of the study (i.e points scored) whereas deep learning method do not have any impact.  

The explanatory variables were selected based on the absolute correlation values. The three explanatory variables for exam points (top correlated variables, also shown in the plot above)) are student's attitude towards learning (attitude), learning strategy (stra) and surface learning approach (surf). The model based on three dependent variables on exam points has the maximum residual value of 10.9 and median of 0.5. Here, residual value is the remaining value after the predicted value is substracted by the observed value. The model summary showed that attitude is highly significant (Pr=1.93e-08) variable that affects the student's exam points. On the other hand, learning strategy and surface learning are not significant variables (Pr>0.01). 
```{r}
model<-lm(points ~ attitude + stra + surf, data = lrn2014)
summary(model)
```
The summary of the model after removing insignificant variables is as follows.
```{r}
model_sig<-lm(points ~ attitude, data = lrn2014)
summary(model_sig)
```

```{r, fig.width=11, fig.height=9}
par(mfrow = c(2,2))
plot(model_sig, which = c(1,2,5))
```

