# Regression and Model validation

*Describe the work you have done this week and summarize your learning.*

- Describe your work and results clearly. 
- Assume the reader has an introductory course level understanding of writing and reading R code as well as statistical methods
- Assume the reader has no previous knowledge of your data or the more advanced methods you are using  

During the second week of this course, we have been delving deeper into R and statistics. We learned about regression models and the application of R in statistical modeling. The datacamp exercises along with the two embedded videos provided good background on the topics. Chapter three from ["An Introduction to Statistical Learning with Applications in R"](http://www-bcf.usc.edu/~gareth/ISL/) covered in-depth topics in linear regression. 

## RStudio Exercise 2
### Data Wrangling
After going through the study materials, I attempted the RStudio exercise. The first part of the exercise was related to data wrangling where a subset of table was generated from a table with raw data (observations). The R script used to create the table can be found [here](https://github.com/kisun/IODS-project/tree/master/data/create_learning2014.R).

### Analysis
The R script for this part is available [here](https://github.com/kisun/IODS-project/tree/master/data/analysis.R).
```{r}
lrn2014<-read.table("data/learning2014.csv")
str(lrn2014)
dim(lrn2014)
```

```{r, fig.width=11, fig.height=9}
library(GGally)
library(ggplot2)
plot_lrn2014<-ggpairs(lrn2014, mapping = aes(col=gender, alpha = 0.3), lower=list(combo = wrap ("facethist", bins = 20)))
plot_lrn2014
```
```{r}
model<-lm(points ~ attitude + stra + surf, data = lrn2014)
summary(model)
```
```{r}
model_sig<-lm(points ~ attitude, data = lrn2014)
summary(model_sig)
```

```{r, fig.width=11, fig.height=9}
par(mfrow = c(2,2))
plot(model_sig, which = c(1,2,5))
```